---
title: "QuantumCyteWorkflow"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "2025-06-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

download packages
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidyverse)
library(car)
```

These are the packages that I have installed:
install.packages("tidyr")
install.packages("ggplot2")
install.packages("tidyverse")
install.packages("car")

download data sets:
```{r cars}
quantfinal <- read.csv("FinalData.csv")
```

Bar graph of the counts of each variant dectected per method:
```{r pressure, echo=FALSE}
#Rename the variant column for ease of use
quantfinal <- quantfinal %>%
  rename(variant_type = `Variant.seen.0.both..1.Qcyte.only..2.manual.only`)

 #Keep only the values 0, 1, 2
gene_detection_counts <- quantfinal %>%
  filter(variant_type %in% c(0, 1, 2)) %>%
  group_by(Gene, variant_type) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(
    method = case_when(
      variant_type == 0 ~ "Both",
      variant_type == 1 ~ "Qcyte only",
      variant_type == 2 ~ "Manual only"
    )
  )

# Plot grouped bar chart
ggplot(gene_detection_counts, aes(x = Gene, y = count, fill = method)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("Both" = "#F2C75C", 
                               "Qcyte only" = "#3A913F", 
                               "Manual only" = "#54585A")) + 
  labs(title = "Gene Detection by Method",
       x = "Gene",
       y = "Number of Detections",
       fill = "Detection Method") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


From the above Bar Graph, a Z-test was conducted for only the genes that were detected by both QuantumCyte and Manual:
```{r}
# List of genes you're interested in
genes_of_interest <- c("APC", "BRAF", "BRCA1", "GNAS", "IDH1", "KRAS", "PMS2")

# Initialize results list
z_results <- data.frame(
  Gene = character(),
  Z_statistic = numeric(),
  P_value = numeric(),
  stringsAsFactors = FALSE
)

# Loop through each gene
for (gene in genes_of_interest) {
  # Filter and remove NA pairs
  gene_data <- quantfinal %>%
    filter(Gene == gene) %>%
    select(Qcyte.AF, Manual.VAF) %>%
    na.omit()
  
  if (nrow(gene_data) > 1) {  # need at least 2 observations for SD
    # Paired values
    paired_Qcyte.AF <- gene_data$Qcyte.AF
    paired_Manual.VAF <- gene_data$Manual.VAF
    n <- length(paired_Qcyte.AF)
    
    # Z-test calculation
    mean_diff <- mean(paired_Qcyte.AF) - mean(paired_Manual.VAF)
    sd_diff <- sd(paired_Qcyte.AF - paired_Manual.VAF)
    z <- mean_diff / (sd_diff / sqrt(n))
    p <- 2 * (1 - pnorm(abs(z)))
    
    # Store result
    z_results <- rbind(z_results, data.frame(
      Gene = gene,
      Z_statistic = z,
      P_value = p
    ))
  } else {
    # Add NA if not enough data
    z_results <- rbind(z_results, data.frame(
      Gene = gene,
      Z_statistic = NA,
      P_value = NA
    ))
  }
}

# Print results
print(z_results)

```


Boxplots of the genes with significant Z-tests (the other genes didn't have enough samples):
```{r}
#Graph for APC Variant
data_apc <- quantfinal %>% 
  filter(Gene == "APC")

boxplot(data_apc$Manual.VAF, data_apc$Qcyte.AF,
        names = c("Manual.VAF", "Qcyte.AF"),
        main = "Box Plot for APC Variant Qcyte.AF vs Manual.VAF",
        ylab = "VAF",
        col = c("#A4D65E", "#F8E08E"),
        border = c("#3A913F", "#F2C75C"))

points(x = rep(1, nrow(data_apc)), 
       y = data_apc$Manual.VAF, 
       pch = 19, col = "black")

# For Qcyte.AF (right box at x=2):
points(x = rep(2, nrow(data_apc)), 
       y = data_apc$Qcyte.AF, 
       pch = 19, col = "black")

# Connect the paired points (from the same row) with lines.
segments(x0 = rep(1, nrow(data_apc)), 
         y0 = data_apc$Manual.VAF, 
         x1 = rep(2, nrow(data_apc)), 
         y1 = data_apc$Qcyte.AF, 
         col = "black")

#Graph for BRAF Variant
data_braf <- quantfinal %>% 
  filter(Gene == "BRAF")

boxplot(data_braf$Manual.VAF, data_braf$Qcyte.AF,
        names = c("Manual.VAF", "Qcyte.AF"),
        main = "Box Plot for BRAF Gene Qcyte.AF vs Manual.VAF",
        ylab = "VAF",
        col = c("#A4D65E", "#F8E08E"),
        border = c("#3A913F", "#F2C75C"))

points(x = rep(1, nrow(data_braf)), 
       y = data_braf$Manual.VAF, 
       pch = 19, col = "black")

# For Qcyte.AF (right box at x=2):
points(x = rep(2, nrow(data_braf)), 
       y = data_braf$Qcyte.AF, 
       pch = 19, col = "black")

# Connect the paired points (from the same row) with lines.
segments(x0 = rep(1, nrow(data_braf)), 
         y0 = data_braf$Manual.VAF, 
         x1 = rep(2, nrow(data_braf)), 
         y1 = data_braf$Qcyte.AF, 
         col = "black")


#Graph for KRAS Variant
data_kras <- quantfinal %>% 
  filter(Gene == "KRAS")

boxplot(data_kras$Manual.VAF, data_kras$Qcyte.AF,
        names = c("Manual.VAF", "Qcyte.AF"),
        main = "Box Plot for KRAS Variant Qcyte.AF vs Manual.VAF",
        ylab = "VAF",
        col = c("#A4D65E", "#F8E08E"),
        border = c("#3A913F", "#F2C75C"))

points(x = rep(1, nrow(data_kras)), 
       y = data_kras$Manual.VAF, 
       pch = 19, col = "black")

# For Qcyte.AF (right box at x=2):
points(x = rep(2, nrow(data_kras)), 
       y = data_kras$Qcyte.AF, 
       pch = 19, col = "black")

# Connect the paired points (from the same row) with lines.
segments(x0 = rep(1, nrow(data_kras)), 
         y0 = data_kras$Manual.VAF, 
         x1 = rep(2, nrow(data_kras)), 
         y1 = data_kras$Qcyte.AF, 
         col = "black")


#Graph for PMS2 Variant
data_pms2 <- quantfinal %>% 
  filter(Gene == "PMS2")

boxplot(data_pms2$Manual.VAF, data_pms2$Qcyte.AF,
        names = c("Manual.VAF", "Qcyte.AF"),
        main = "Box Plot for PMS2 Gene Qcyte.AF vs Manual.VAF",
        ylab = "VAF",
        col = c("#A4D65E", "#F8E08E"),
        border = c("#3A913F", "#F2C75C"))

points(x = rep(1, nrow(data_pms2)), 
       y = data_pms2$Manual.VAF, 
       pch = 19, col = "black")

# For Qcyte.AF (right box at x=2):
points(x = rep(2, nrow(data_pms2)), 
       y = data_pms2$Qcyte.AF, 
       pch = 19, col = "black")

# Connect the paired points (from the same row) with lines.
segments(x0 = rep(1, nrow(data_pms2)), 
         y0 = data_pms2$Manual.VAF, 
         x1 = rep(2, nrow(data_pms2)), 
         y1 = data_pms2$Qcyte.AF, 
         col = "black")


```

